The importance of measurement uncertainty in ecological management

### 2018-01-04 appendix/appendix_harvestfirst.Rmd
discount at 0.95
WIP

```{r}
options(mc.cores=parallel::detectCores())
```

@@

## ASSUMES harvest takes place *before* recruitment, f(x-h), not after.
escapement_policy <- function(x) pmax(x - S_star,0)  
x0 <- K/6
Tmax = 20
det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0

@@

sims <- 
  list(msy = msy_policy, 
       pgy = pgy_policy,
       det = escapement_policy) %>% 
  map_df(det_sim, 
         f, x0, Tmax, 
         .id = "method") 
sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1)

@@

policies <- data.frame(
  det = index(pmax(f(states,0) - S_star,0), actions), # assumes obs->recruit->harvest, f(x_t) - h_t
 # det = index(pmax(states - S_star,0),     actions), # assumes obs->harvest->recruit, f(x_t - h_t)
  msy = index(states * F_MSY,               actions),
  pgy = index(states * F_PGY,               actions))
```

@@

```{r}
meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.15), 
                    sigma_m = c(0, 0.1, 0.15),
                    stringsAsFactors = FALSE) %>%
        mutate(scenario  = as.character(1:length(sigma_m)))
log_dir <- "appendix_alphas_harvestfirst" # Store the computed solution files here
meta
```

@@

```{r eval=TRUE}
dir.create(log_dir)

@@

```{r}
meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% 
  mutate(scenario = as.character(scenario)) %>%
  left_join(meta) %>% 
  arrange(scenario)
alphas <- alphas_from_log(meta, log_dir)
```

@@

```{r static_sims}
Tmax <- 100
x0 <- which.min(abs(K/6 - states))
reps <- 100

static_sims <- 
 map_dfr(models, function(m){
            do_sim <- function(policy) sim_pomdp(
                        m$transition, m$observation, m$reward, discount, 
                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
            map_dfr(policies, do_sim, .id = "method")
          }, .id = "scenario")

@@

```{r pomdp_sims}
# slower!
unif_prior <- rep(1, length(states)) / length(states)
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
                       reps = reps)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario")
```

@@

write_csv(sims, "appendix_harvestfirst_files/sims.csv")
```

@@

```{r}
sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

@@

write_csv(policy_table, "appendix_files/policy_table.csv")
```

@@

pomdp_policies <- df %>% 
  select(-value) %>% 
  rename(uniform = policy) %>% 
  gather(policy, harvest, -state) %>%
  mutate(state = states[state], 
         harvest = actions[harvest]) %>%
  mutate(escapement = state-harvest) %>%
  select(state, policy, harvest, escapement)

@@

>>>>

```{r}
options(mc.cores=parallel::detectCores())
log_dir <- "../data/observe_harvest_recruit" # Store the computed solution files here

@@

## ASSUMES harvest takes place *before* recruitment, f(x-h), not after.
escapement_policy <- function(x) pmax(x - S_star,0)

x0 <- K/6
Tmax = 20
do_det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0

@@

det_sims <- 
  list(msy = msy_policy, 
       pgy = pgy_policy,
       det = escapement_policy) %>% 
  map_df(do_det_sim, 
         f, x0, Tmax, 
         .id = "method")

With no stochasticity, MSY leads to the same long-term stock size as under the constant escapement rule, but takes longer to get there.  This level is essentially $B_{MSY}$, though because the model considered here implements events in the order: *observe, harvest, recruit*; rather than *observe, recruit, harvest*, we see the stock at the pre-harvest size of $B_{MSY} + H_{MSY}$ (that is, $K/2 + rK/4$).   More conservative rules, such as a harvest set to 80% of MSY result in faster recovery of the stock than under MSY, but slower than under constant escapement.  Due to the reduced maximum harvest rate, such rules lead to stock returning to a value higher than $B_${MSY}$.

```{r}
det_sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1)

@@

policies <- data.frame(
#  det = index(pmax(f(states,0) - S_star,0), actions), # obs,recruit,harv, f(x_t) - h_t
  det = index(escapement_policy(states),     actions), # obs,harv,recruit, f(x_t - h_t)
  msy = index(msy_policy(states),               actions),
  pgy = index(pgy_policy(states),               actions))
```

@@

```{r}
meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.15), 
                    sigma_m = c(0, 0.1, 0.15),
                    stringsAsFactors = FALSE) %>%
        mutate(scenario  = as.character(1:length(sigma_m)))
meta
```

@@

```{r eval=FALSE}
dir.create(log_dir)

@@

```{r}
meta <- meta_from_log(data.frame(model="gs", discount=discount), log_dir) %>% 
  mutate(scenario = as.character(scenario)) %>%
  left_join(meta) %>% 
  arrange(scenario)
alphas <- alphas_from_log(meta, log_dir)
```

@@

```{r static_sims}
set.seed(12345)

Tmax <- 100
x0 <- which.min(abs(K/6 - states))
reps <- 100
static_sims <- 
 map_dfr(models, function(m){
            do_sim <- function(policy) sim_pomdp(
                        m$transition, m$observation, m$reward, discount, 
                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
            map_dfr(policies, do_sim, .id = "method")
          }, .id = "scenario")

@@

```{r pomdp_sims}
set.seed(12345)

unif_prior <- rep(1, length(states)) / length(states)
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
                       reps = reps)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario")
```

@@

sims <- sims %>% select(time, state, rep, method, sigma_m, sigma_g)

sim_col_types <- paste0(gsub("^(\\w).*", "\\1", sapply(unname(sims), class)), collapse = "")
write_csv(sims, file.path(log_dir, "sims.csv"))
```

@@

```{r}
sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)

sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
  facet_grid(sigma_m ~ sigma_g, 
             labeller = label_bquote(sigma[m] == .(sigma_m),
                                     sigma[g] == .(sigma_g))) + 
  coord_cartesian(ylim = c(0, 1))

@@

@@

pomdp_policies <- df %>% 
  select(state, uniform = policy, low, medium, high) %>% 
  gather(policy, harvest, -state) %>%
  mutate(state = states[state], 
         harvest = actions[harvest]) %>%
  mutate(escapement = state-harvest) %>%
  select(state, policy, harvest, escapement)

all_policies <- bind_rows(policy_table, pomdp_policies)

write_csv(all_policies, file.path(log_dir, "all_policies.csv"))

@@

```{r}
all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "msy")) %>%
  mutate(policy = fct_recode(policy,
                             "MSY" = "msy",
                             "Constant escapement" = "det",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

```{r}
all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "msy")) %>%
  mutate(policy = fct_recode(policy,
                             "MSY" = "msy",
                             "Constant escapement" = "det",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) %>%
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

### 2018-01-04 manuscripts/manuscript.Rmd
WIP

bibliography: citations.bib
output: rticles::elsevier_article
---

@@

- Replicate simulations

Moderate / modest sigma_g, sigma_m

- non-pomdp policies

- pomdp policies

- historical data.

```{r}
```

>>>>

bibliography: citations.bib
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column

---

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
library(tidyverse)
library(gridExtra)
library(ggthemes)
colours <- ptol_pal()(4)
berkeley_colors <- c("#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")
```

```{r suggested_libs}
library(hrbrthemes)
library(extrafont)
##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
extrafont::loadfonts()
hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
```

```{r}
log_dir <- "../data/observe_harvest_recruit" # solution files here
```

@@

<!--
PGY, MSY, Det, MDP (sigma_g), POMDP (sigma_g, sigma_m)
-->

### Replicate simulations

Figure 1 shows average fish biomass

```{r fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty. Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement, maximum sustainable yield (MSY), and the porposed partially observed Markov decision process (POMDP) method. Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.  Environmental stochasticity is fixed a standard deviation of 0.15 in each panel.  Carrying capacity K normalized to 1, r = 0.75. Additional environmental noise levels shown in supplement."}
sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnn")
sims %>%
  filter(method != "pgy", sigma_g == "0.15") %>%
  mutate(method = fct_recode(method, 
                             "Constant escapement" = "det",
                             "MSY" = "msy",
                             "POMDP" = "pomdp")) %>%
  group_by(time, method, sigma_m) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
  coord_cartesian(ylim = c(0, 1)) + 
  theme(legend.position = "bottom") + 
  ylab("Mean biomass")
```

POMDP policies depend on the prior belief.

```{r}
all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "msy")) %>%
  mutate(policy = fct_recode(policy,
                             "MSY" = "msy",
                             "Constant escapement" = "det",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line(lwd=1)  + 
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
```

### Comparisons to historical data

```{r eval=FALSE}
sims %>%
  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = c("black", "grey", berkeley_colors) ) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```
